---
title: "assignment"
author: "Paula Iller"
date: "2024-05-24"
output: html_document
---

```{r}
suppressPackageStartupMessages({
  library(GenomicRanges)
  library(AnnotationHub)
  library(ensembldb)
  library(InteractionSet)
})
```

# Download data 

```{r}
options(timeout = 10000)
# p300 peaks in cells:
download.file("https://www.encodeproject.org/files/ENCFF776TFX/@@download/ENCFF776TFX.bed.gz", "p300.narrowPeak.gz")
download.file("https://ethz-ins.org/content/hg38.SalviatoDistalEnhancerTargets.GR.rds", "salviato.rds")

peaks <- rtracklayer::import("p300.narrowPeak.gz")
seqlevelsStyle(peaks) = "ensembl"
targetSalviato <- readRDS("salviato.rds")
seqlevelsStyle(targetSalviato) = "ensembl"

hubCache(AnnotationHub())
ah <- AnnotationHub()
ensdb <- ah[["AH95744"]]

annotatedPeaks <- epiwraps::annotateRegions(peaks, ensdb)

head(annotatedPeaks)
```

```{r}
# Extend the TSS regions between 2.5kb and 10kb
p1 <- promoters(targetSalviato, 2500, 10000)
peaks$overlapsP1 <- overlapsAny(peaks, p1)
peaksP1 <- peaks[peaks$overlapsP1]
peaksP1 
```
```{r}
# Extend the TSS regions over 10kb
p2 <- promoters(targetSalviato, 10000)
peaks$overlapsP2 <- overlapsAny(peaks, p2)
peaksP2 <- peaks[peaks$overlapsP2]
peaksP2 
```
```

```{r}
# find overlaps
o1 <- findOverlaps(peaksP1, targetSalviato)
head(o1)
mcols(peaksP1)[from(o1),"target"] <- targetSalviato[to(o1)]$target
mcols(peaksP1)$target <- CharacterList(mcols(peaksP1)$target)
TSS1 <- round((length(peaksP1) / length(targetSalviato)), digits = 4)


o2 <- findOverlaps(peaksP2, targetSalviato)
head(o2)
mcols(peaksP2)[from(o2),"target"] <- targetSalviato[to(o2)]$target
mcols(peaksP2)$target <- CharacterList(mcols(peaksP2)$target)
TSS2 <- round((length(peaksP2) / length(targetSalviato)), digits = 4)
```

```{r}
paste("Of the genes that are between 2.5 and 10kb from the nearest TSS,", TSS1 ,"% form an interaction with that nearest gene. Of the genes that are more than 10kb away from the nearest TSS, ",TSS2 ,"% form an interaction with that nearest gene.")
```


